---
title: "EDS_assignment_5"
author: "M.Kinneen"
date: "2023-02-18"
output:
  pdf_document: default
  toc: true
  number_sections: true
---


## 1. Environment set up 

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggridges)
```
\pagebreak

## 2. Data read

Read in and check data.

```{r read}
sisco_data <- read.csv("./data/siscowet.csv", header = TRUE)

#have a look
head(sisco_data)
summary(sisco_data)

#check for NA
sapply(sisco_data, function(x) sum(is.na(x)))
```
Sex and age columns have large amounts of NA values. Also one in weight.
\pagebreak

## 3. Initial exploratory plots

```{r init expl., fig.dim=c(3,3)}
univar_exploratory <- function(df){
  columns<-c(colnames(df)) #create list of columns
  for(i in 1:length(columns)){   #Loop through plotting hist of each
  var<- df[,i]   #select var for plotting
  title<-paste(columns[i])
  if(is.numeric(var)==TRUE){   #if var is numeric, plot, otherwise pass 
    hist(var,
         main = title)
  }else{
    print(paste0(title,"(column ",i,") ", "is a non-numeric column"))
  }
  }
}

#length_weight plot to check outliers
ggplot(sisco_data,aes(x=len,y = log(wgt)))+
  geom_point()+
  stat_smooth(method = "lm")
```

```{r, fig.dim=c(3,3)}
univar_exploratory(df = sisco_data)

```

PNLdepth is approximately normal, median ~ 70.Mesh has right skew, median 2 - 3. Possibly log transform? Fish ID approxiamtely uniform, irrelevant for analysis.Age has right skew, median age of ~ 11. Length is normal, median of 500. Weight is heavy right skew, lagre max value likely outlier (confirmed by length-weight plot). filtering needed

\pagebreak




## 4. Clean and filter data

Weight less skewed though still non-normal. Do not do more filtering as larger fish
will be underrepresented in data.
```{r clean}
sisco_data_cleaned <- sisco_data
sisco_data_cleaned<- sisco_data%>%
  filter(wgt <= 7000)%>% #filter max value from wgt
  mutate(mesh_log = log(mesh)) #log transform mesh
```

```{r, fig.dim=c(3,3)}
#Repeat plots
univar_exploratory(df = sisco_data_cleaned)

```

\pagebreak

## 5. Exploratory plot

Plot showing the length distribution of fish across sites.
```{r explor.}
exploratory<-ggplot(sisco_data_cleaned,aes(x=len))+
  geom_histogram()+
  xlab("length")+
  labs(title = "Length distribution for siscowet lake trout")

exploratory
ggsave("./images/exploratory.jpg") #save image


#save image

```
\pagebreak

## 6. Expository plot

Fish length distributions across sites.

```{r expos., results = FALSE}

palette<- brewer.pal(4,"Set2") #colorblinf friendly paeltte
site_names<- unique(sisco_data_cleaned$locID) #site names for labels
samples<-table(sisco_data_cleaned$locID) # count data for lavbels

expository_plot<-
  #Define data
  ggplot(sisco_data_cleaned,aes(x = len, y = locID, alpha =0.05, fill = locID))+
  #Define geometry (histograms)
  geom_density_ridges(stat = "binline", scale = 1, size = 0.01)+
  #Overlay curve
  geom_density_ridges(scale = 1, size = 0.01)+
    scale_fill_manual(values = palette)+
  #Set axis breaks
  scale_x_continuous(breaks = seq(200,800,150))+
  #Add title
  labs(title = substitute(paste("Length distributions of Siscowet Lake trout ",
                     italic("(Salvelinus namaycush)"))),
       #Data source caption
       caption = paste("Data source: FSAdata","\n",
                        "github.com/fishR-Core-Team/FSAdata"))+
  xlab("Fish length (mm)")+
  ylab("Sampling site")+
  #Simple theme
  theme_classic()+
  #Custom theme - remove y axis title, adjust title position 
  theme(
    axis.text.y  = element_blank(),
    legend.position = "none",
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 7),
    plot.caption = element_text(size = 5)
  )+
  #Add annotations for site name and number of samples.
  annotate("text",x = 750, y = 1.5, label = site_names[2],colour=palette[1])+
  annotate("text",x = 750, y = 1.35, label = paste0("n = ",samples[1])
           ,colour=palette[1],fontface = 3,size = 3)+
  annotate("text",x = 750, y = 2.5, label = site_names[1],colour=palette[2])+
   annotate("text",x = 750, y = 2.35, label = paste0("n = ",samples[2])
           ,colour=palette[2],fontface = 3,size = 3)+
  annotate("text",x = 750, y = 3.5, label = site_names[4],colour=palette[3])+
   annotate("text",x = 750, y = 3.35, label = paste0("n = ",samples[3])
           ,colour=palette[3],fontface = 3,size = 3)+
  annotate("text",x = 750, y = 4.5, label = site_names[3],colour=palette[4])+
   annotate("text",x = 750, y = 4.35, label = paste0("n = ",samples[4])
           ,colour=palette[4],fontface = 3,size = 3)


```

```{r, results = FALSE}
expository_plot
ggsave("./images/expository.jpg")

```


